Process-based phenotypes: an application to the ontology of cell physiology

### 2011-11-30 paper/drinking.tex

In computer science, an ontology is a specification of a
conceptualization of a domain of knowledge \cite{Gruber1995,
  Guarino1998}.  Ontologies commonly distinguish between {\em classes}
and {\em individuals} within a domain of knowledge. A class is an
entity that can have {\em instances}, while individuals are entities
that cannot be instantiated. Examples of individuals include the
Eiffel tower or the 2009 Ironman Triathlon in Hawaii, while examples
of classes include {\em Tower} or {\em Triathlon}. The Eiffel tower
can be an instance of the class {\em Tower}, and the 2009 Ironman
Triathlon an instance of {\em Triathlon}.  The meaning of classes is
specified by stating what must be true of their instances.

In addition to classes and individuals, ontologies often include {\em
  relations}. Relations are the ``the glue that holds things together,
the primary constituents of the facts that go to make up reality''
\cite{Barwise1989}.

@@

Eventually, a set of {\em primitive terms} remains that are not
further defined. Using only the primitive terms, we can construct
complex sentences. Based on the intended meaning of the primitive
terms, we consider some of these sentences true and some of them false
in our domain. We select some of the true sentences as {\em axioms}
which provide the core of our ontology. Ideally, the axioms are chosen
so that all true sentences in the domain we intend to represent follow
by means of logical deduction from the axioms. More commonly, however,
only {\em some aspects} of the intended meaning are formally
represented while other aspects are omitted either due to limitations
in language expressivity or due to their irrelevance to the problem
for which an ontology is developed.

Based on the axioms and definitions, we can use deduction to infer
statements that logically follow from the axioms.  The process of
automatically deducing sentences from axioms is called {\em automated
  reasoning}. Automated reasoning allows users of an ontology to carry
our key activities: verifying the ontology's consistency, inferring
hidden knowledge and thereby performing powerful queries.  An ontology
is inconsistent if {\em every} statement can be inferred from the
ontology. In particular, an ontology is inconsistent if a statement
$\phi$ and its negation $\neg \phi$ can be inferred from the
ontology's axioms. Classes in an ontology may be unsatisfiable: a
class $C$ is unsatisfiable, if it is not possible that the class has
instances. Unsatisfiable classes are commonly the result of a
contradictory class definition. That a class is unsatisfiable is often
a special kind of unintended consequence that can be drawn from an
ontology.

>>>>

In computer science, an ontology is a specification of a
conceptualization of a domain of knowledge \cite{Gruber1995,
  Guarino1998}.  Ontologies commonly distinguish between {\em classes}
(also called {\em concepts}, {\em categories} or {\em universals}) and
{\em individuals} within a domain of knowledge. A class is an entity
that can have {\em instances}, while individuals are entities that
cannot be instantiated \cite{Herre2006}. Examples of individuals
include the Eiffel tower or the 2009 Ironman Triathlon in Hawaii,
while examples of classes include {\em Tower} or {\em Triathlon}. The
Eiffel tower can be an instance of the class {\em Tower}, and the 2009
Ironman Triathlon an instance of {\em Triathlon}.  The meaning of
classes is specified by stating what must be true of their instances.

In addition to classes and individuals, ontologies often include {\em
  relations}. Relations hold between entities, they are the ``the glue
that holds things together, the primary constituents of the facts that
go to make up reality'' \cite{Barwise1989}.

@@

Eventually, a set of {\em primitive terms} remains that are not
further defined. Following the axiomatic method \cite{Hilbert1918},
using only the primitive terms, we can construct complex
sentences. Based on the intended meaning of the primitive terms, we
consider some of these sentences true and some of them false in our
domain. We select some of the true sentences as {\em axioms} which
provide the core of our ontology. Ideally, the axioms are chosen so
that all true sentences in the domain we intend to represent follow by
means of logical deduction from the axioms. More commonly, however,
only {\em some aspects} of the intended meaning are formally
represented while other aspects are omitted either due to limitations
in language expressivity or due to their irrelevance to the problem
for which an ontology is developed.

Based on the axioms and definitions, we can use deduction to infer
statements that logically follow from the axioms.  The process of
automatically deducing sentences from axioms is called {\em automated
  reasoning}. Automated reasoning allows users of an ontology to carry
our key activities: verifying the ontology's consistency, inferring
hidden knowledge and thereby performing powerful queries.  An ontology
is formally inconsistent if there is a statement $\phi$ such that
$\phi$ and its negation $\neg \phi$ can be inferred from the
ontology's axioms.  If an ontology is formally inconsistent, {\em
  every} statement can be inferred from the ontology.

We assume that the classes of an ontology are
satisfiable, i.e. that they have instances

Classes in an ontology may be unsatisfiable: a class $C$ is
unsatisfiable, if it is not possible that the class has
instances. Unsatisfiable classes are commonly the result of a
contradictory class definition. That a class is unsatisfiable is often
a special kind of unintended consequence that can be drawn from an
ontology.
